---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

### Resample covariates such that they are 
```{r}
library(raster)
refraster = readRDS("data/from_tf_mapping/ppp_20km_af_2020_20220929.rds")
fls = list.files(path="data/from_tf_mapping/", pattern=".*.rds$", full.names=T)
covars = vector('list', length(fls))
for (i in seq_along(fls)) {
  cat("i =", i, ", ")
  covars[[i]] <- raster::resample(readRDS(fls[[i]]), refraster, method='bilinear')
}
for (i in seq_along(fls)) {
  covname = gsub(".*mapping/|_2022.*", "", fls[[i]])
  writeRaster(covars[[i]], paste0("data/covars/", covname, "_", tstamp(), ".tif"))
}
# Malaria
afss_adm0 <- readRDS("data/africa_sub_Sahara_adm0_shp.rds")
fls = list.files(path="data/", pattern=".*Pf.*.tif$", full.names=T)
covars2 = vector('list', length(fls))
for (i in seq_along(fls)) { # first file duplicate
  # rst = raster(fls[[i]])
  # covars2[[i]] = tidy_raster(rst=raster(fls[[i]]), extent=afss_adm0, ref=refraster)
  # covname = gsub(".*data/|.tif", "", fls[[i]])
  covname = gsub(".*data/|_([0-9]+).tif", "", fls[[i]])
  year = gsub('.*_([0-9]+).*','\\1', fls[[i]])
  fname = paste0("data/covars/", covname, "_af_", year, "_", tstamp(), ".tif")
  print(fname)
  writeRaster(covars2[[i]], fname)
}
```

### Improved sanitation estimates
These were downloaded later
Access to improved sanitation facilities and relying on unimproved sanitation facilities. Although there is a separate variable, access to non-piped improved sanitation facilities, we did not include as it is a subset of the variable, access to improved sanitation facilities, which has easier interpretation. 
```{r}
library(raster)
refraster = readRDS("data/from_tf_mapping/ppp_20km_af_2020_20220929.rds")
fls = list.files(path="data/", pattern="S_UNIMP_PERCENT|S_IMP_PERCENT", full.names=T)

covars = vector('list', length(fls))
for (i in seq_along(fls)) {
  cat("i =", i, ", ")
  rst = raster(fls[i])
  covars[[i]] <- tidy_raster(rst, func="mean", ref=refraster)
  # covars[[i]] <- raster::resample(readRDS(fls[[i]]), refraster, method='bilinear')
}
for (i in seq_along(fls)) {
  # covname = gsub(".*data/|_([0-9]+).tif", "", fls[[i]])
  covname = gsub(".*data/|_([0-9]+)_([0-9]+)|_([0-9]+)_Y2020.*.TIF", "", fls[[i]])
  year = gsub('^data.*_MEAN_|_Y2020.*', "", fls[[i]])
  fname = paste0("data/covars/", covname, "_20km_af_", year, "_", tstamp(), ".tif")
  print(fname)
  writeRaster(covars[[i]], fname)
}
```

### RasterBrick of covariates by year
Creates RasterBrick of covariates

```{r}
elev = raster("data/covars/elevation_20km_africa.rds_20230314.tif")
motor_traveltime_healthcase = raster("data/covars/motorized_travel_time_to_healthcare_20km_af_20230314.tif")
walk_traveltime_healthcase = raster("data/covars/walking_travel_time_to_healthcare_20km_af_20230314.tif")
traveltime_city = raster("data/covars/travel_time_cities_20km_af_20230314.tif")
distance_to_water = raster("data/covars/distance_to_water_20km_af_20230314.tif")
## load the covars files by year

years = 2000:2020
for (yr in years) {
  pat = paste0(".*af_", yr, ".*tif$")
  # fls = list.files(path="data/covars/", pattern=".*af_2000.*tif$", full.names=T)
  fls = list.files(path="data/covars/", pattern=pat, full.names=T)
  # fn = list.files(path="data/covars/", pattern=".*af_2000.*tif$")
  # the following covariates are absent for 2018-2020 and we used the 2017 instead
  if (yr >= 2018){
    fls = c(fls[1:2]
   , "data/covars/annual_mean_temperature_20km_af_2017_20230314.tif"   
   , "data/covars/annual_rainfall_20km_af_2017_20230314.tif"           
   , "data/covars/improved_water_20km_af_2017_20230314.tif"            
   , "data/covars/open_defecation_20km_af_2017_20230314.tif"           
   , "data/covars/piped_sanitation_20km_af_2017_20230314.tif"          
   , "data/covars/piped_water_20km_af_2017_20230314.tif"
   , fls[3]
   ,"data/covars/prev_HIV_adults_20km_af_2017_20230314.tif"           
   , "data/covars/prev_stunting_20km_af_2017_20230314.tif"             
   , "data/covars/surface_water_20km_af_2017_20230314.tif"             
   , "data/covars/underweight_20km_af_2017_20230314.tif"               
   , "data/covars/wasting_20km_af_2017_20230314.tif")
  }
  cat("year =", yr, ", #fls =", length(fls), "\n")
  
  covars = vector('list', (length(fls)+5))
  for (i in seq_along(fls)){
    covars[[i]] = raster(fls[[i]])
  }
  
  covars[[i+1]] = elev
  covars[[i+2]] = motor_traveltime_healthcase
  covars[[i+3]] = walk_traveltime_healthcase
  covars[[i+4]] = traveltime_city
  covars[[i+5]] = distance_to_water
  # covstack = raster::stack(covars)
  # 
  nms = lapply(fls, function(x) gsub('data/covars/|_20230314.tif$', "", x))
  # names(covstack) = c(unlist(nms), "elevation", "motor_traveltime_healthcare", "walk_traveltime_healthcare", "traveltime_city")
  # ## add the covars that are not dependent by 
  # terra::writeRaster(covstack, paste0("data/covars/covars_", yr, ".tif"), overwrite=TRUE)
  # r = terra::rast(paste0("data/covars/covars_", yr, ".tif"))
  
  covbrick = raster::brick(covars)
  names(covbrick)= c(unlist(nms), "elevation", "motor_traveltime_healthcare", "walk_traveltime_healthcare", "traveltime_city", "distance_to_water")
  # native raster format (.grd) retains the names for each RasterLayer while .tif does not
  writeRaster(covbrick, filename=paste0("data/covars/covars_brick_", yr, "_", tstamp(), ".grd"), format="raster", overwrite=TRUE, options=c("INTERLEAVE=BAND","COMPRESS=LZW"))
}
# infile = brick(paste0("data/covars/covars_", yr, ".grd"))
# infile
# infile[[2]]
```

