---
title: "Probability of Occurrence of Invasive Non-Typhoidal Salmonella"
author: "Jong-Hoon Kim"
date: '2023-02-18'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
devtools::load_all()
```

### Clean data
```{r}
library(readxl)
library(dplyr)
library(magrittr)
dat <- read_xlsx("data/iNTS_occurrence_in_sub-Saharan_Africa.xlsx", 
               sheet = "iNTS",
               col_types = c("text", "numeric", "text", "text", "date",
                             "date", rep("text",3), "numeric", rep("text",
                             8), rep("numeric", 2), "text"))

dat$YEAR_BEGIN <- as.integer(dat$YEAR_BEGIN)
dat$YEAR_END <- as.integer(dat$YEAR_END)
# the either of two exists, then 
dat %<>% filter(INCLUDE == "YES") %>%
  mutate(YEAR_BEGIN = ifelse(is.na(YEAR_BEGIN), YEAR_END, YEAR_BEGIN))
dat %<>% filter(INCLUDE == "YES") %>% 
  mutate(YEAR_END = ifelse(is.na(YEAR_END), YEAR_BEGIN, YEAR_END))
                                                              
# location, year, diagnostic methods are needed
dat %<>% filter(!is.na(LATITUDE), !is.na(LONGITUDE),
                !is.na(YEAR_BEGIN), !is.na(YEAR_END),
                !is.na(DIAGNOSTIC_METHODS), INCLUDE == "YES")

dat$YEAR_BEGIN <- as.numeric(dat$YEAR_BEGIN)
dat$YEAR_END <- as.numeric(dat$YEAR_END)
names(dat) <- tolower(names(dat))

dat %<>% filter(diagnostic_methods != "Urine culture")
#
saveRDS(dat, paste0("outputs/occ_", tstamp(), ".rds"))
data.table::fwrite(dat, paste0("outputs/occ_", tstamp(), ".csv"))
```

### Descriptive statistics
```{r}
library(dplyr)
# source("R/util.R")
# d <- readRDS("outputs/occ_20230223.rds")
d <- data.table::fread(paste0("outputs/occ_20230317.csv"))

# unique references
d_distinct <- distinct(d, study_info, pub_year, .keep_all=T)
(n_study <- nrow(distinct(d, study_info, pub_year)))
(n_country <- length(unique(d_distinct$country)))
# study years vary
(min(d$year_begin, na.rm=T))
(max(d$year_end, na.rm=T))

# number of studies by county
d_distinct %>% 
  dplyr::group_by(country) %>%
  dplyr::summarize(count=n()) -> dco
# over half of the reports came from 
dco$percent <- 100*dco$count/sum(dco$count)
dco[order(dco$percent, decreasing = T),]
cumsum(dco[order(dco$percent, decreasing = T),]$percent)

# number of cases by country
d_distinct %>% 
  dplyr::group_by(country) %>%
  dplyr::summarize(total_case=sum(num_case)) -> d2  

sum(d2$total_case)
d2$percent <- 100*d2$total_case/sum(d2$total_case)
d2[order(d2$percent,decreasing=T),]
cumsum(d2[order(d2$percent,decreasing=T),]$percent)

# diagnostic method
unique(d_distinct$diagnostic_methods)
blood <- grep("blood", d_distinct$diagnostic_methods, ignore.case=T, value=T)
CSF <- grep("CSF|cerebrospinal", d_distinct$diagnostic_methods, ignore.case=T, value=T)

pleural <- grep("pleural", d_distinct$diagnostic_methods, ignore.case=T, value=T)
peritoneal <- grep("peritoneal", d_distinct$diagnostic_methods, ignore.case=T, value=T)

length(blood)
length(blood)/nrow(d_distinct)
length(CSF)
length(CSF)/nrow(d_distinct)
length(pleural)
length(pleural)/nrow(d_distinct)
length(peritoneal)
length(peritoneal)/nrow(d_distinct)
# hospital type

d_area <- d_distinct[is.na(d_distinct$catchment_area), ]
nrow(d_area)/nrow(d_distinct)
unique(d_distinct$hospital_type)
primary = grep("primary", d_distinct$hospital_type, ignore.case=T, value=T)
tertiary = grep("teaching|referral|reference|tertiary", d_distinct$hospital_type, ignore.case=T, value=T)
area = d_distinct$hospital_type[d_distinct$hospital_type == "NA"]
length(primary) / nrow(d_distinct)
length(tertiary) / nrow(d_distinct)
length(area) / nrow(d_distinct)

```


### Fig 2A. Number of reports by year and country
```{r}
# d <- data.table::fread(paste0("outputs/occ_20230223.csv"))
d <- data.table::fread(paste0("outputs/occ_20230317.csv"))
d <- distinct(d, study_info, pub_year, .keep_all=T)
library(dplyr)
names(d)
d %>% 
  dplyr::group_by(country) %>%
  dplyr::summarize(count=n()) -> dcntry

d %>% 
  dplyr::group_by(country, pub_year) %>%
  dplyr::summarize(count=n(), percent = count/sum(count)) -> dcoyr

d %>%
  dplyr::group_by(pub_year) %>%
  dplyr::summarize(count = n()) -> d_tot

d_tot$country <- "Total"
dcoyr <- bind_rows(dcoyr, d_tot)

cntry_ordered <- unique(dcoyr$country)
dcoyr$country <- factor(dcoyr$country, levels = rev(cntry_ordered))

cntry_labels <- cntry_ordered
cntry_labels[2] <- "CAR" # abbreviation for better displaying

library(ggplot2)

p <- ggplot(data=dcoyr) + 
  geom_point(aes(x=pub_year, y=country, size=count)) +
  # geom_point(aes(x = year_pub, y = country, size = count_lab)) +
  scale_size_continuous(limits = c(1, 50)) +
  scale_x_continuous(breaks=seq(2000, 2020, by=5), limits = c(2000,2020)) + 
  scale_y_discrete(breaks = cntry_ordered, labels = cntry_labels) + 
  labs(x="Year of publication", y="", size="No. of reports" ) +
  theme_bw() +
  theme(legend.position = "bottom")+
  theme(text=element_text(size=14))   

print(p)

ffac <- 2
ggsave(paste0("plots/reports_cntry_yr_", tstamp(hour=T), ".png"), p, width=3.4*ffac, height=2.7*ffac, units="in")
```

### Fig 2B. Number of cases by year and country
1. Ignore the when the cases are not 
2. Select unique studies by checking author, year of publication, year of surveillance, and number of case.
3. The number of cases evenly distributed across years


#### Create yearly dataset
```{r}
# dataet by year
d <- data.table::fread(paste0("outputs/occ_20230317.csv"))
library(dplyr)
# if end > begin  then create additional YEAR_END
d2 = d[d$year_end == d$year_begin,]
d2$year = d2$year_end

for(i in 1:nrow(d)) {
  yeardiff = d$year_end[i] - d$year_begin[i]
  if (yeardiff > 0) {
    avgcase = d$num_case [i] / yeardiff
    while (yeardiff > 0){
      drow = d[i,]
      drow$year = d$year_end[i] - yeardiff
      # avgcase decimal points because cases numbers
      # are evenly distributed across data collection years
      drow$num_case = round(avgcase)
      d2 = rbind(d2, drow)
      yeardiff = yeardiff - 1
    }
  }
}
data.table::fwrite(d2, paste0("outputs/occ_yearly_", tstamp(), ".csv"))
```

#### Plot case vs. year
```{r}
# d <- data.table::fread(paste0("outputs/occ_20230316.csv"))
library(dplyr)
d2 = data.table::fread(paste0("outputs/occ_yearly_20230317.csv"))

# over half of the reports came from 
d2 %>% 
  dplyr::group_by(country) %>%
  dplyr::summarize(case=sum(num_case)) -> dcntry

d2 %>% 
  dplyr::group_by(country, year_begin) %>%
  dplyr::summarize(case=sum(num_case)) -> dcy

d2 %>%
  dplyr::group_by(year_begin) %>%
  dplyr::summarize(case=sum(num_case)) -> dtot

dtot$country <- "Total"
dcoyr <- bind_rows(dcy, dtot)

cntry_ordered <- unique(dcoyr$country)
dcoyr$country <- factor(dcoyr$country, levels = rev(cntry_ordered))

cntry_labels <- cntry_ordered
cntry_labels[2] <- "CAR" # abbreviation for better displaying

# tiff("figs/reports_co_yr.tiff", units="in", width=5.4, height=7, res=300)
library(ggplot2)

p <- ggplot(data=dcoyr) + 
  geom_point(aes(x=year_begin, y=country, size=case)) +
  # geom_point(aes(x = year_pub, y = country, size = count_lab)) +
  # scale_size_continuous(limits = c(1, 50)) +
  scale_x_continuous(breaks=seq(1990, 2020, by=5), limits = c(1988,2020)) + 
  scale_y_discrete(breaks = cntry_ordered, labels = cntry_labels) + 
  labs(x="Year of data collection", y="", size="No. of cases" ) +
  theme_bw() +
  theme(legend.position = "bottom")+
  theme(text=element_text(size=14))  
  
p

ffac <- 2

ggsave(paste0("plots/cases_cntry_yr_", tstamp(hour=T), ".png"), p, width=3.4*ffac, height=2.7*ffac, units="in")
```


### Create grid occurrence dataset
This step involves assigning grid cells to each of the cases based on the exponential decay if cases occurred in a certain hospital and random grids if the subnational area was known.
```{r}
d = read.csv("outputs/occ_yearly_20230317.csv")
shplist = readRDS(paste0("outputs/occ_shapelist_20230317.rds"))
# shplist = get_shapelist(dat=d, fuzzy=T)
# check polygons using the script below
# saveRDS(shplist, paste0("outputs/occ_shapelist_", tstamp(), ".rds"))
# Use Check polygons script to make sure that all shapefiles are intact 
# for some areas, we don't have shapefiles and 
# I conducted get_shapelist iteratively to make sure that I get the shapelist all filled in 
ndata = 100 
dists = c(50, 100, 200)
for (ds in dists){
  occdata = vector("list", ndata)
  for (i in 1:length(occdata)){
    cat("i =" , i, "\n")
    set.seed(i)
    occdata[[i]] = create_occ_data(dat=d, dist3=ds, shapelist=shplist)
  }
  saveRDS(occdata, 
          paste0("outputs/occ_dataset_dist3_", ds, "_", tstamp(), ".rds"))
}
```


### Create grid occurrence raster
This is simply to create 20 km by 20 km raster that reflects the number or presence of iNTS 
```{r}
library(raster)
# d = read.csv("outputs/occ_yearly_20230317.csv")
# dataset = readRDS("outputs/occ_dataset_20230317T111754.rds")
fls = list.files("outputs/", "occ_dataset_dist3_.*_20230318.rds", full.names=TRUE)
# dataset = readRDS("outputs/occ_dataset_dist3_100_20230318.rds")

# this is just a template and values will will be replaced
rst = raster::raster("data/covars/ppp_20km_af_2000_20230314.tif")
# shplist = readRDS("outputs/occ_shapelist_20230317.rds")
for (fl in fls) {
  dataset = readRDS(fl)
  # cat("seed =", i, "\n")
  d = dataset[[1]]
  rstocc = create_occ_raster(dat=d, raster=rst)
  dist = gsub("outputs/occ_dataset_dist3_|_20230318.rds", "", fl)
  saveRDS(rstocc, paste0("outputs/occ_raster_cumul_dist_", dist, "_", tstamp(hour=T), ".rds"))
}

# d = dataset[[1]]
# rstocc = create_occ_raster(dat=d, raster=rst)
# saveRDS(rstocc, paste0("outputs/occ_raster_cumul_", tstamp(hour=T), ".rds"))


# for different seeds
for (i in 1:length(dataset)) {
  cat("seed =", i, "\n")
  d = dataset[[i]]
  rstocc = create_occ_raster(dat=d, raster=rst)
  saveRDS(rstocc, paste0("outputs/occ_raster_cumul_seed_", i, "_", tstamp(hour=T), ".rds"))
}

## create yearly raster
# rst <- raster::raster("data/covars/ppp_20km_af_2000_20230314.tif")
years = 2000:2020
occrst_yearly_lst = vector("list", length(years))
names(occrst_yearly_lst) <- years
set.seed(42)
for (yr in years) {
  cat("yr =", yr, "\n")
  dtemp = d[d$year == yr,]
  if (nrow(dtemp) > 0) {
    occrst_yearly_lst[[as.character(yr)]] = 
    create_occ_raster(dat=dtemp, raster=rst)
  }
}
saveRDS(occrst_yearly_lst, paste0("outputs/occ_raster_yearly_", tstamp(hour=T), ".rds"))
```


### Grid descriptive statistics 
Number of grids have reported one or more occurrence of iNTS out of total grids
```{r}
occrst <- readRDS(paste0("outputs/occ_raster_cumul_20230317T11.rds"))
occrst_occ <- occrst[occrst>=1]
occrst_occ2 <- occrst[occrst>=2]
# proportion of grid cells more than 2 occurrences
(length(occrst_occ))
(length(occrst_occ) / length(occrst))
fls = list.files("outputs", pattern="occ_raster_cumul_seed", full.names=T)
prop = c()
case = c()
for (fl in fls){
  occ = readRDS(fl)
  occ_over_1 = occ[occ>=1]
  prop = c(prop, length(occ_over_1) / length(occ))
  # print(prop)
  case = c(case, length(occ_over_1))
}
summary(prop)
summary(case)
quantile(prop, probs=c(0.025, 0.5, 0.975))
quantile(case, probs=c(0.025, 0.5, 0.975))

occrstyr <- readRDS(paste0("outputs/occ_raster_yearly_20230317T11.rds"))
years = 2000:2020
for (yr in years) {
  rst = occrstyr[[as.character(yr)]]
  totlen = length(rst)
  len = length(rst[rst>=1])
  cat(length(rst[rst>=1]), ", ", len/totlen, "\n")
}
```


### Fig S3: Cumulative iNTS occurrence on grids
Number of occurrence with different colors on grids
```{r}
fls = list.files("outputs/", "occ_raster_cumul_dist_.*_20230318T09.rds", full.names=TRUE)

for (fl in fls) {
  occrst = readRDS(fl)
  plt = occ_plot(raster=occrst, plot_yesno=FALSE)
  dist = gsub("outputs/occ_raster_cumul_dist_|_20230318T09.rds", "", fl)

  rat <- map_ratio(occrst)
  ggsave(paste0("plots/occ_cumul_num_grid_", dist, "_",
              tstamp(hour=T), ".png"), plt, width=7.4,
       height=7.4*rat, units="in")
}
# presence or absence plot
for (fl in fls) {
  occrst = readRDS(fl)
  plt = occ_plot(raster=occrst, plot_yesno=TRUE)
  dist = gsub("outputs/occ_raster_cumul_dist_|_20230318T09.rds", "", fl)

  rat <- map_ratio(occrst)
  ggsave(paste0("plots/occ_cumul_yesno_grid_", dist, "_",
              tstamp(hour=T), ".png"), plt, width=7.4,
       height=7.4*rat, units="in")
}
```

### Fig 3: Cumulative iNTS occurrence on grids
```{r}
occrst = readRDS("outputs/occ_raster_cumul_dist_100_20230318T09.rds")
plt = occ_plot(raster=occrst, plot_yesno=TRUE, add_zoom=TRUE)

# part that is magnified colored differently
e = extent(28,40,-24,-8) # extent to zoom in QECH, Blantyre, Malawi
rcrop <- crop(occrst, e, snap="out")
rcrop <- crop(rcrop, afss_adm0, snap="out")
rcrop <- mask(rcrop, afss_adm0) # without this, sea becomes grey..
rpts_crop <- as.data.frame(rasterToPoints(rcrop))
colnames(rpts_crop) <- c("lon","lat","num_occ")

occrst_mask <- mask(occrst, afss_adm0) # without this, sea becomes grey..
rptsdf <- as.data.frame(rasterToPoints(occrst_mask))
colnames(rptsdf) <- c("lon","lat","num_occ")

shape0 <- afss_adm0
shape1 <- afss_adm1

rptsdf$num_occ <- ifelse(rptsdf$num_occ == 0, NA, rptsdf$num_occ)

library(ggplot2)
library(RColorBrewer)

mypal <- brewer.pal(9, "YlOrBr")

p <- ggplot(rptsdf) +
  geom_polygon(data=shape0, aes(long, lat, group = group),
               fill="grey95", inherit.aes = FALSE)+
  geom_raster(aes(lon, lat, fill=num_occ), alpha=0.8) +
  scale_fill_gradientn(colors=mypal,"Number of\niNTS occurrence")+
  geom_raster(data=rpts_crop, aes(lon, lat), fill="steelblue", alpha=0.4, inherit.aes = FALSE) +
  geom_path(data=shape0, aes(long, lat, group=group),
            color="black", linewidth=0.8, inherit.aes=FALSE) +
  geom_path(data=shape1, aes(long, lat, group=group),
            color="black", linewidth=0.4, inherit.aes=FALSE) +
  coord_equal() +
  theme_map() +
  theme(legend.title = element_text(size=12),
        legend.text = element_text(size=12))
print(p)
# 
rat <- map_ratio(occrst)
ggsave(paste0("plots/occ_cumul_num_grid_", yr, "_", tstamp(hour=T), ".png"), p, width=7.4, height=7.4*rat, units="in")
```

#### Fig 3: Cumulative iNTS occurrence on grids (Yes or No)
```{r}
library(raster)
afss_adm1 <- readRDS("data/africa_sub_Sahara_adm1_shp.rds")
afss_adm0 <- readRDS("data/africa_sub_Sahara_adm0_shp.rds")

lst = readRDS("outputs/occ_raster_yearly_20230317T11.rds")
yr = 2010
occrst = lst[[as.character(yr)]]
# seed <- 1
# occrst = readRDS("outputs/occ_raster_cumul_seed_1_20230317T11.rds")
# part that is magnified colored differently
e = extent(28,40,-24,-8) # extent to zoom in QECH, Blantyre, Malawi
rcrop <- crop(occrst, e, snap="out")
rcrop <- crop(rcrop, afss_adm0, snap="out")
rcrop <- mask(rcrop, afss_adm0) # without this, sea becomes grey..
rpts_crop <- as.data.frame(rasterToPoints(rcrop))
colnames(rpts_crop) <- c("lon", "lat", "num_occ")

rst_res <- mask(occrst, afss_adm0) # without this, sea becomes grey..
rpts <- rasterToPoints(rst_res)
rptsdf <- as.data.frame(rpts)
colnames(rptsdf) <- c("lon", "lat", "num_occ")
shape0 <- afss_adm0
shape1 <- afss_adm1

# rptsdf$num_occ <- ifelse(rptsdf$num_occ == 0, NA, rptsdf$num_occ)
rptsdf$num_occ <- ifelse(rptsdf$num_occ == 0, "No", "Yes")
rptsdf$num_occ <- as.factor(rptsdf$num_occ)
library(ggplot2)
library(RColorBrewer)

p <- ggplot(rptsdf) +
  geom_polygon(data=shape0, aes(long, lat, group = group),
               fill=NA, inherit.aes = FALSE)+
  geom_raster(aes(lon, lat, fill=num_occ), alpha=0.8) +
  scale_fill_manual(values=c("Yes"="brown","No"="white"),
                    labels=c("Yes"="Yes","No"="No"),
                    "iNTS occurrence")+
  geom_raster(data=rpts_crop, aes(lon, lat), fill="steelblue",
              alpha=0.4, inherit.aes = FALSE) +
  geom_path(data=shape0, aes(long, lat, group=group),
            color="black", linewidth=0.8, inherit.aes=FALSE) +
  geom_path(data=shape1, aes(long, lat, group=group),
            color="black", linewidth=0.4, inherit.aes=FALSE) +
  coord_equal() +
  theme_map() +
  theme(legend.title = element_text(size=12))

print(p)

rat <- map_ratio(occrst)
# ggsave(paste0("plots/occ_cumul_yesno_grid_", yr, "_", tstamp(hour=T), ".png"), p, width=7.4, height=7.4*rat, units="in")
ggsave(paste0("plots/occ_2010_yesno_grid_", yr, "_", tstamp(hour=T), ".png"), p, width=7.4, height=7.4*rat, units="in")
```


#### Fig 3: iNTS occurrence on grids - small section

Yes or No
```{r}
occrst = readRDS("outputs/occ_raster_cumul_seed_1_20230317T11.rds")
yr = "cumul"
# lst = readRDS("outputs/occ_raster_yearly_20230317T11.rds")
# yr = 2010
# occrst = lst[[as.character(yr)]]

e = extent(28,40,-24,-8) # extent to zoom in QECH, Blantyre, Malawi
afss_adm1_crop = crop(afss_adm1, e, snap="out")
afss_adm0_crop = crop(afss_adm0, e, snap="out")
occrst_mask <- mask(occrst, afss_adm1) # without this, sea becomes grey..
rst_crop = crop(occrst_mask, e, snap="out")
rpts <- rasterToPoints(rst_crop)
rptsdf <- as.data.frame(rpts)
colnames(rptsdf) <- c("lon", "lat", "num_occ")
shape0 <- afss_adm0_crop
shape1 <- afss_adm1_crop

rptsdf$num_occ <- ifelse(rptsdf$num_occ == 0, "No", "Yes")
rptsdf$num_occ <- as.factor(rptsdf$num_occ)
library(ggplot2)
library(RColorBrewer)

p <- ggplot(rptsdf) +
  geom_polygon(data=shape0, aes(long, lat, group = group),
               fill=NA, inherit.aes = FALSE)+
  geom_raster(aes(lon, lat, fill=num_occ), alpha=0.8) +
  scale_fill_manual(values=c("Yes"="brown","No"="white"),
                    "iNTS occurrence")+
  # geom_raster(data=rpts_crop, aes(lon, lat), fill="steelblue",
  #             alpha=0.4, inherit.aes = FALSE) +
  geom_path(data=shape0, aes(long, lat, group=group),
            color="black", linewidth=0.8, inherit.aes=FALSE) +
  geom_path(data=shape1, aes(long, lat, group=group),
            color="black", linewidth=0.4, inherit.aes=FALSE) +
  coord_equal() +
  theme_map() +
  guides(fill='none')+
  theme(legend.title = element_text(size=12))

print(p)

rat <- map_ratio(occrst)
ggsave(paste0("plots/occ_", yr, "_yesno_grid_smaller_", tstamp(hour=T), ".png"), p, width=7.4, height=7.4*rat, units="in")

# mypal <- brewer.pal(9, "YlOrBr")
# 
# p <- ggplot(rptsdf) +
#   geom_polygon(data = shape0, aes(long, lat, group = group),
#                fill = "grey90", inherit.aes = FALSE)+
#   geom_raster(aes(lon, lat, fill=num_occ+1), show.legend=F) +
#   # geom_raster(aes(lon, lat), fill="steelblue", alpha=0.4) +
#   scale_fill_gradientn(colors = mypal, "iNTS occurrence", 
#                        trans = 'log10') +
#   geom_path(data = shape0, aes(long, lat, group = group),
#             color = "black", linewidth=0.8, inherit.aes = FALSE) +
#   geom_path(data = shape1, aes(long, lat, group = group),
#             color = "black", linewidth=0.4, inherit.aes = FALSE) +
#   coord_equal() +
#   theme_map() +
#   # guides(fill=FALSE)+
#   theme(legend.title = element_text(size=12),
#         legend.text = element_text(size=12))
# 
# print(p)


```



```{r}
e = extent(28,40,-24,-8) # extent to zoom in QECH, Blantyre, Malawi
afss_adm1_crop = crop(afss_adm1, e, snap="out")
afss_adm0_crop = crop(afss_adm0, e, snap="out")
occrst_mask <- mask(occrst, afss_adm1) # without this, sea becomes grey..
rst_crop = crop(occrst_mask, e, snap="out")
rpts <- rasterToPoints(rst_crop)
rptsdf <- as.data.frame(rpts)
colnames(rptsdf) <- c("lon", "lat", "num_occ")
shape0 <- afss_adm0_crop
shape1 <- afss_adm1_crop

rptsdf$num_occ <- ifelse(rptsdf$num_occ == 0, NA, rptsdf$num_occ)
library(ggplot2)
library(RColorBrewer)

mypal <- brewer.pal(9, "YlOrBr")

p <- ggplot(rptsdf) +
  geom_polygon(data = shape0, aes(long, lat, group = group),
               fill = "grey90", inherit.aes = FALSE)+
  geom_raster(aes(lon, lat, fill=num_occ+1), show.legend=F) +
  # geom_raster(aes(lon, lat), fill="steelblue", alpha=0.4) +
  scale_fill_gradientn(colors = mypal, "Number of iNTS occurrence", 
                       trans = 'log10') +
  geom_path(data = shape0, aes(long, lat, group = group),
            color = "black", linewidth=0.8, inherit.aes = FALSE) +
  geom_path(data = shape1, aes(long, lat, group = group),
            color = "black", linewidth=0.4, inherit.aes = FALSE) +
  coord_equal() +
  theme_map() +
  # guides(fill=FALSE)+
  theme(legend.title = element_text(size=12),
        legend.text = element_text(size=12))

print(p)

rat <- map_ratio(rst)
ggsave(paste0("plots/occ_cumul_grid_smaller_", tstamp(), ".png"), p, width=7.4, height=7.4*rat, units="in")

```



#### Fig 3-1: Yearly iNTS occurrence on grids (Yes or No)
```{r}
library(raster)
afss_adm1 <- readRDS("data/africa_sub_Sahara_adm1_shp.rds")
afss_adm0 <- readRDS("data/africa_sub_Sahara_adm0_shp.rds")
# seed <- 1
lst = readRDS("outputs/occ_raster_yearly_20230317T09.rds")
yr = 2000
occrst = lst[[as.character(yr)]]
# 
# part that is magnified colored differently
e = extent(28,40,-24,-8) # extent to zoom in QECH, Blantyre, Malawi
rcrop <- crop(occrst, e, snap="out")
rcrop <- crop(rcrop, afss_adm0, snap="out")
rcrop <- mask(rcrop, afss_adm0) # without this, sea becomes grey..
rpts_crop <- as.data.frame(rasterToPoints(rcrop))
colnames(rpts_crop) <- c("lon", "lat", "num_occ")

rst_res <- mask(occrst, afss_adm0) # without this, sea becomes grey..
rpts <- rasterToPoints(rst_res)
rptsdf <- as.data.frame(rpts)
colnames(rptsdf) <- c("lon", "lat", "num_occ")
shape0 <- afss_adm0
shape1 <- afss_adm1

# rptsdf$num_occ <- ifelse(rptsdf$num_occ == 0, NA, rptsdf$num_occ)
rptsdf$num_occ <- ifelse(rptsdf$num_occ == 0, "No", "Yes")
rptsdf$num_occ <- as.factor(rptsdf$num_occ)
library(ggplot2)
library(RColorBrewer)

p <- ggplot(rptsdf) +
  geom_polygon(data=shape0, aes(long, lat, group = group),
               fill=NA, inherit.aes = FALSE)+
  geom_raster(aes(lon, lat, fill=num_occ), alpha=0.8) +
  scale_fill_manual(values=c("Yes"="brown","No"="white"), "Grids of\niNTS occurrence")+
  geom_raster(data=rpts_crop, aes(lon, lat), fill="steelblue", alpha=0.4, inherit.aes = FALSE) +
  geom_path(data=shape0, aes(long, lat, group=group),
            color="black", linewidth=0.8, inherit.aes=FALSE) +
  geom_path(data=shape1, aes(long, lat, group=group),
            color="black", linewidth=0.4, inherit.aes=FALSE) +
  coord_equal() +
  theme_map() +
  theme(legend.title = element_text(size=12),
        legend.text = element_blank())
# print(p)

rat <- map_ratio(occrst)
ggsave(paste0("plots/occ_yesno_grid_", yr, "_", tstamp(hour=T), ".png"), p, width=7.4, height=7.4*rat, units="in")
```





### Typical occurrence grid cells for a single hospital 

```{r}
refstudy = c(pub_year = 2016, year_begin=2006, year_end=2006, 
location="Queen Elizabeth Central Hospital", country="Malawi")
dref = filter(d, pub_year == refstudy[["pub_year"]], year_begin==refstudy[["year_begin"]], year_end==refstudy[["year_end"]], country == refstudy[["country"]])
rst <- readRDS("data/elevation_20km_africa.rds")
# plot(rst)
rstocc <- rst
rstocc[] <- 0

place <- dref$location
ncase <- as.numeric(dref$num_case)

lonlat <- cbind(dref$longitude, dref$latitude)
for(i in 1:ncase) {
  cat("i =", i, "\n")
  occ = pick_random_point(x=lonlat, r=0.01)
  cell = raster::extract(rstocc, occ, small=T, cellnumber=T)
  ids = unique(as.data.frame(cell)$cell)
  rstocc[ids] <- rstocc[ids] + 1
}
library(raster)

afss_adm1 <- readRDS("data/africa_sub_Sahara_adm1_shp.rds")
afss_adm0 <- readRDS("data/africa_sub_Sahara_adm0_shp.rds")
# smaller extent to zoom in the area
e = extent(31, 39, -19, -11)
afss_adm1_crop = crop(afss_adm1, e, snap="out")
afss_adm0_crop = crop(afss_adm0, e, snap="out")
rstocc_mask <- mask(rstocc, afss_adm1) # without this, sea becomes grey..
rstocc_mask = crop(rstocc_mask, e, snap="out")

rpts <- rasterToPoints(rstocc_mask)
rptsdf <- as.data.frame(rpts)

colnames(rptsdf) <- c("lon", "lat", "num_occ")
shape <- afss_adm1_crop
shape0 <- afss_adm0_crop

rptsdf$num_occ <- ifelse(rptsdf$num_occ == 0, NA, rptsdf$num_occ)
library(ggplot2)
library(RColorBrewer)

mypal <- brewer.pal(9, "YlOrBr")

p <- ggplot(rptsdf) +
  geom_polygon(data = shape, aes(long, lat, group = group),
                fill = "grey80", inherit.aes = FALSE)+
  geom_raster(aes(lon, lat, fill=num_occ)) +
  scale_fill_gradientn(colors = mypal, "Number of iNTS occurrence") +
  geom_path(data = shape, aes(long, lat, group = group),
             color = "black", linewidth = 0.5, inherit.aes = FALSE) +
  geom_path(data = shape0, aes(long, lat, group = group),
             color = "black", linewidth = 1.0, inherit.aes = FALSE) +
  coord_equal() +
  theme_map() +
  theme(legend.title = element_text(size=12),
        legend.text = element_text(size=12))
p

rat <- map_ratio(rst)
ggsave(paste0("plots/occ_grid_QECH_", tstamp(), ".png"), p, width=7.4, height=7.4*rat, units="in")
```



#### Fig 2C-1: iNTS occurrence on grids (Yes or No)
```{r}
library(raster)
afss_adm1 <- readRDS("data/africa_sub_Sahara_adm1_shp.rds")
afss_adm0 <- readRDS("data/africa_sub_Sahara_adm0_shp.rds")
# seed <- 1
lst <- readRDS("outputs/occ_raster_list_20230313T19.rds")
occrst = lst[[1]]
# part that is magnified colored differently
e = extent(28,40,-24,-8) # extent to zoom in QECH, Blantyre, Malawi
rcrop <- crop(occrst, e, snap="out")
rcrop <- crop(rcrop, afss_adm0, snap="out")
rcrop <- mask(rcrop, afss_adm0) # without this, sea becomes grey..
rpts_crop <- as.data.frame(rasterToPoints(rcrop))
colnames(rpts_crop) <- c("lon", "lat", "num_occ")

rst_res <- mask(occrst, afss_adm0) # without this, sea becomes grey..
rpts <- rasterToPoints(rst_res)
rptsdf <- as.data.frame(rpts)
colnames(rptsdf) <- c("lon", "lat", "num_occ")
shape0 <- afss_adm0
shape1 <- afss_adm1

# rptsdf$num_occ <- ifelse(rptsdf$num_occ == 0, NA, rptsdf$num_occ)
rptsdf$num_occ <- ifelse(rptsdf$num_occ == 0, "No", "Yes")
rptsdf$num_occ <- as.factor(rptsdf$num_occ)
library(ggplot2)
library(RColorBrewer)

p <- ggplot(rptsdf) +
  geom_polygon(data=shape0, aes(long, lat, group = group),
               fill=NA, inherit.aes = FALSE)+
  geom_raster(aes(lon, lat, fill=num_occ), alpha=0.8) +
  scale_fill_manual(values=c("Yes"="brown","No"="white"), "Grids of\niNTS occurrence")+
  geom_raster(data=rpts_crop, aes(lon, lat), fill="steelblue", alpha=0.4, inherit.aes = FALSE) +
  geom_path(data=shape0, aes(long, lat, group=group),
            color="black", linewidth=0.8, inherit.aes=FALSE) +
  geom_path(data=shape1, aes(long, lat, group=group),
            color="black", linewidth=0.4, inherit.aes=FALSE) +
  coord_equal() +
  theme_map() +
  theme(legend.title = element_text(size=12),
        legend.text = element_blank())
print(p)

rat <- map_ratio(occrst)
ggsave(paste0("plots/occ_grid_yesno_", tstamp(hour=T,minute=T), ".png"), p, width=7.4, height=7.4*rat, units="in")
```


#### Fig 2D-1: iNTS occurrence on grids - small section

```{r}
e = extent(28,40,-24,-8) # extent to zoom in QECH, Blantyre, Malawi
afss_adm1_crop = crop(afss_adm1, e, snap="out")
afss_adm0_crop = crop(afss_adm0, e, snap="out")
occrst_mask <- mask(occrst, afss_adm1) # without this, sea becomes grey..
rst_crop = crop(occrst_mask, e, snap="out")
rpts <- rasterToPoints(rst_crop)
rptsdf <- as.data.frame(rpts)
colnames(rptsdf) <- c("lon", "lat", "num_occ")
shape0 <- afss_adm0_crop
shape1 <- afss_adm1_crop

rptsdf$num_occ <- ifelse(rptsdf$num_occ == 0, "No", "Yes")
rptsdf$num_occ <- as.factor(rptsdf$num_occ)
library(ggplot2)
library(RColorBrewer)

p <- ggplot(rptsdf) +
  geom_polygon(data=shape0, aes(long, lat, group = group),
               fill=NA, inherit.aes = FALSE)+
  geom_raster(aes(lon, lat, fill=num_occ), alpha=0.8) +
  scale_fill_manual(values=c("Yes"="brown", "No"="white"), guide="none")+
  geom_path(data=shape0, aes(long, lat, group=group),
            color="black", linewidth=0.8, inherit.aes=FALSE) +
  geom_path(data=shape1, aes(long, lat, group=group),
            color="black", linewidth=0.4, inherit.aes=FALSE) +
  coord_equal() +
  theme_map()
  
print(p)

rat <- map_ratio(occrst)
ggsave(paste0("plots/occ_grid_yesno_smaller_", tstamp(hour=T,minute=T), ".png"), p, width=7.4, height=7.4*rat, units="in")

```


### Occurrence grids - descriptive statistics
How many of the grid cells have occurrence? 

```{r}
occ <- readRDS("outputs/occ_20230223.rds")
source("R/util.R") # to use clean_country_names function
source("R/get_shape.R")
source("R/get_occ_ncell.R")
library(raster)
elev <- readRDS("data/elevation_20km_africa.rds")
(sum(!is.na(elev[])))
afss_adm0 <- readRDS("data/africa_sub_Sahara_adm0_shp.rds")
# elev_afss <- readRDS("outputs/elevation_20km_africa.rds")
elev_afss_crop <- raster::crop(elev, extent(afss_adm0), snap = "out")
elev_afss_crop_mask <- raster::mask(elev_afss_crop, mask = afss_adm0)

(noNA_grids <- sum(!is.na(elev_afss_crop_mask[])))

rst <- elev_afss_crop_mask

cellnumbers <- rep(NA, nrow(occ))
for (i in 1:nrow(occ)){
  oc <- data.frame(X=occ$long[i], Y=occ$lat[i])
  cells <- raster::extract(rst, oc, df = TRUE, cellnumbers = TRUE)
  cellnumbers[i] <- unlist(cells$cells)
}
cellfreq <- as.data.frame(table(cellnumbers))
nrow(cellfreq) / noNA_grids
cellfreq <- cellfreq[order(cellfreq$Freq, decreasing = T),]
dupl <- cellfreq[cellfreq$Freq > 1,]

# change the resoltuion
clist <- list()
ft <- 1:5
cellnumbers <- data.frame(matrix(NA, nrow = nrow(occ), ncol=length(ft)))
ngrids <- rep(NA, length(ft))

for (j in 1:length(ft)) {
  rst_res <- raster::aggregate(rst, fact = ft[j], fun = mean)
  # rst_res <- raster::disaggregate(rst, fact = ft[j], fun = mean)
  ngrids[j] <- sum(!is.na(rst_res[]))
  for (i in 1:nrow(occ)){
    oc <- data.frame(X=occ$long[i], Y=occ$lat[i])
    cells <- raster::extract(rst_res, oc, df = TRUE, cellnumbers = TRUE)
    # if(!is.na(unlist(cells$cells)))
    cellnumbers[i, j] <- unlist(cells$cells)
  }
  cellfreq <- as.data.frame(table(cellnumbers[,j]))
  cellfreq <- cellfreq[order(cellfreq$Freq, decreasing = T),]
  clist[[j]] <- cellfreq 
}
# number of grids 
sapply(clist, function(z) nrow(z)) / ngrids

clist <- list()
ft <- c(2, 4)
cellnumbers <- data.frame(matrix(NA, nrow = nrow(occ), ncol=length(ft)))
ngrids <- rep(NA, length(ft))

for (j in 1:length(ft)) {
  rst_res <- raster::disaggregate(rst, fact = ft[j], fun = mean)
  ngrids[j] <- sum(!is.na(rst_res[]))
  for (i in 1:nrow(occ)){
    oc <- data.frame(X=occ$long[i], Y=occ$lat[i])
    cells <- raster::extract(rst_res, oc, df = TRUE, cellnumbers = TRUE)
    cellnumbers[i, j] <- unlist(cells$cells)
  }
  cellfreq <- as.data.frame(table(cellnumbers[,j]))
  cellfreq <- cellfreq[order(cellfreq$Freq, decreasing = T),]
  clist[[j]] <- cellfreq 
}
# number of grids 
sapply(clist, function(z) nrow(z)) / ngrids
```

Number of grids in Africa is `r noNA_grids`

