---
title: "Probability of Occurrence of Invasive Non-Typhoidal Salmonella"
author: "Jong-Hoon Kim"
date: '2023-02-18'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
devtools::load_all()
```

## Boosted regression trees 

### Create occurrence and absence data
```{r}
library(raster)
library(dplyr)
# occ = readRDS("outputs/occ_dataset_20230314T1231.rds")
# occfull = readRDS("outputs/occ_dataset_20230316T173539.rds")
# occfull = readRDS("outputs/occ_dataset_20230317T111754.rds")
occfull = readRDS("outputs/occ_dataset_dist3_200_20230318.rds")
refrst = raster("data/covars/motorized_travel_time_to_healthcare_20km_af_20230314.tif")
# go through data year by year such that we match the covariates by year and location
for (i in 15) {
  for (j in 16:20) {
    cat("i =", i, ", j =",  j, "\n")
    set.seed(j)
    data_brt = create_occ_abs_data(occdata=occfull[[i]], refraster=refrst)
    saveRDS(data_brt, paste0("outputs/data_dist3_200_brt_occ_", i, "_abs_", j, "_", tstamp(),".rds"))
  }
}
```

### Correlation
```{r}
cov = readRDS("outputs/name_cov_baseline.rds")
occid=1 
absid=1
fls=list.files("outputs", "^data_brt_occ_([0-9]+)_abs_([0-9]+)_202303.*.rds$",full.names=T)
fid = grepl(paste0("occ_",occid,"_abs_",absid,"_"), fls, fixed=T)
sum(fid)
d = readRDS(fls[fid])
d = d[,cov]

names(d) = c("Pf incidence rate","Pf parasite rate","Improved water","Open defecation","Piped sanitation","Piped water","HIV","Stunting", "Surface water","Underweight","Wasting")

library(Hmisc)
cormat = rcorr(as.matrix(d), type="pearson")
corr_r = as.data.frame(round(cormat$r,2))
corr_p = as.data.frame(round(cormat$P,2))

data.table::fwrite(corr_r, paste0("outputs/corr_r_",tstamp(),".csv"), row.names=TRUE)
data.table::fwrite(corr_p, paste0("outputs/corr_p_",tstamp(),".csv"),row.names=T)
```

### Model run
```{r}
library(raster)
library(dismo)
library(scales)
library(ggplot2)
library(dplyr)

# brtdat = readRDS(paste0("outputs/data_brt_20230317T175517.rds"))
# brtdat = readRDS(paste0("outputs/data_dist3_50_brt_occ_1_abs_1_20230322.rds"))

brtdat = readRDS(paste0("outputs/data_dist3_200_brt_occ_1_abs_1_20230322.rds"))
# varnames = readRDS("outputs/name_cov_baseline.rds")
# variables were selected after looking at the variable response curve. In particular, HIV, Pf incidence rate, underweight has positive association (risk factors), improved water and piped sanitation have negative association (protective factors) on March 17, 2023
varnames = c("X202206_Global_Pf_Incidence_Rate",
             "improved_water_20km",
             "piped_sanitation_20km",
             "prev_HIV_adults_20km","underweight_20km")    

brtdat = brtdat[,c("yesno",varnames)]
# yr = 2020
# covpred = brick(paste0("data/covars/covars_brick_", yr, "_20230316.grd"))
# covpred = covpred[[varnames]]

hyperparams = expand.grid(tc=4:6, lr=c(0.01,0.05), bf=seq(0.5,0.8,0.1))

ipp = 1 # iterations per parameter
niter = nrow(hyperparams)
brt_fits <- vector("list", niter*ipp)
# brt_pred <- vector("list", niter*ipp)
# set.seed(42)

# brtdat = brtdat[sample.int(nrow(brtdat), round(nrow(brtdat)/2)),]
# date <- tstamp()
for (i in 1:niter) {
  # brt_fits_by_hyper <- vector("list", ipp)
  for(j in 1:ipp){
    cat("i =", i, ", j =", j, ", time =", tstamp(hour=T, minute=T, second=T), "\n")
    set.seed(j)
    # fit_brt <- gbm.step(data = brtdat, 
    #                     gbm.x = c(2:(ncol(brtdat)-6)), 
    #                     gbm.y = 1,
    #                     family = "bernoulli",
    #                     tree.complexity = 5,
    #                     learning.rate = 0.01,
    #                     bag.fraction = 0.7,
    #                     n.folds = 10,
    #                     silent = TRUE)
    
    fit_brt <- gbm.step(data = brtdat,
                        gbm.x = c(2:ncol(brtdat)),
                        gbm.y = 1,
                        family = "bernoulli",
                        tree.complexity = hyperparams$tc[i],
                        learning.rate = hyperparams$lr[i],
                        bag.fraction = hyperparams$bf[i],
                        silent = TRUE, 
                        plot.main = FALSE)
    
    brt_fits[[(i-1)*ipp+j]] <- fit_brt
    # pred_brt <- predict(covpred,
    #                     fit_brt,
    #                     n.trees = fit_brt$gbm.call$best.trees,
    #                     type = "response")
    
    # brt_pred[[(i-1)*ipp+j]] <- pred_brt
    # brt_fits_by_hyper[[j]] = fit_brt
  }

  # saveRDS(brt_fits_by_hyper, 
  #         paste0("outputs/brt_5var_fit_hyper_", i, "_", date, ".rds"))
}

date <- tstamp(hour=T,minute=T,second=T)
saveRDS(brt_fits, paste0("outputs/brt_5var_dist3_200_fit_occ_1_abs_1_", date, ".rds"))
# saveRDS(brt_pred, paste0("outputs/brt_half_pred_", yr, "_", date, ".rds"))
```


### Model run across datasets (100)

```{r}
devtools::load_all(".")
# library(raster)
# library(dismo)
# library(scales)
# library(ggplot2)
# library(dplyr)

sample_id = seq(1, 400, by=5)
i = sample_id[40]
cat("sample =", i, "\n")

datafiles = list.files("outputs/", "^data_brt_occ_([0-9]+)_abs_([0-9]+)_202303.*.rds", full.names=T)

# varnames = readRDS("outputs/name_cov_baseline.rds")
# variables were selected after looking at the variable response curve. In particular, HIV, Pf incidence rate, underweight has positive association (risk factors), improved water and piped sanitation have negative association (protective factors) on March 17, 2023
varnames = c("X202206_Global_Pf_Incidence_Rate",
             "improved_water_20km",
             "piped_sanitation_20km",
             "prev_HIV_adults_20km","underweight_20km")    

hyper_params = expand.grid(tc=4:6, lr=c(0.01,0.05), bf=seq(0.5,0.8,0.1))
niter = nrow(hyper_params)
iter_per_param = 1 # iterations per parameter

brtdat = readRDS(datafiles[i])
brtdat = brtdat[, c("yesno", varnames)]

occ_id = as.integer(gsub("^outputs/data_brt_occ_|_abs.*", "", datafiles[i]))
abs_id = as.integer(gsub("^outputs/data_brt_occ_([0-9]+)_abs_|_2023.*", "", datafiles[i]))

fits = fit_brt(dat=brtdat, niter=niter, iter_per_param=iter_per_param, hyper_params=hyper_params)

saveRDS(fits, 
        paste0("outputs/fit_5var_occ_", occ_id, "_abs_", abs_id, "_", tstamp(), ".rds"))

```


### Figure 4
#### Find best fits and examine variation across BRT runs
```{r}
# find the best boosted regression fit based on the cross-validation error
# fits = readRDS("outputs/brt_5var_fit_hyper_18_20230317.rds")
# fit_fls = list.files("outputs/", "^brt_5var_fit_hyper_([0-9]+)_2023.*.rds$",
#                      full.names=TRUE)
# avg = c()
# for (fits in fit_fls) {
#   cverr = lapply(readRDS(fits), function(x) x$cv.values[length(x$cv.values)])
#   avg = c(avg, mean(unlist(cverr)))
# }
# id = which.min(avg)
# bestfits = readRDS(fit_fls[id])
# id <- find_best_fit(fits)
# bestfit = fits[[id]]
# cverrors <- lapply(fits, function(x) x$cv.values[length(x$cv.values)])
# summary(unlist(cverrors))

fit_fls = list.files("outputs/", "^fit_5var_occ_([0-9]+)_abs_([0-9]+)_2023.*.rds$",
                      full.names=TRUE)
bestfit_ids = rep(NA, length(fit_fls))
bestfits = vector("list", length(fit_fls))

for (i in seq_along(fit_fls)) {
  fits = readRDS(fit_fls[i]) # across hyperparams
  id = find_best_fit(fits)
  bestfit_ids[i] = id
  bestfits[[i]] = fits[[id]]
}


saveRDS(bestfits, paste0("outputs/bestfits_", tstamp(),".rds"))
saveRDS(bestfits[1:15], paste0("outputs/bestfits_1_", tstamp(),".rds"))
saveRDS(bestfits[16:30], paste0("outputs/bestfits_2_", tstamp(),".rds"))

saveRDS(bestfit_ids, paste0("outputs/bestfits_hyperparam_ids_", tstamp(),".rds"))

```

#### Predictions by year
Use the best fit and make predictions by year
```{r}
extract_ids = function(str){
  occ_id = as.integer(gsub("^outputs/fit_5var_occ_|_abs.*", "", str))
  abs_id = as.integer(gsub("^outputs/fit_5var_occ_([0-9]+)_abs_|_2023.*", "", str))
  return(c(occ_id, abs_id))
}

library(raster)# to extract covariates used in the modeling
# varnames = readRDS("outputs/name_cov_baseline.rds")
varnames = c("X202206_Global_Pf_Incidence_Rate",
             "improved_water_20km",
             "piped_sanitation_20km",
             "prev_HIV_adults_20km","underweight_20km")   

for (i in 16:length(bestfits)) {
  years = 2000:2020
  pred_year = vector("list", length(years))
  names(pred_year) = years
  
  ids = extract_ids(str=fit_fls[i])
  
  for (yr in years) {
    cat("i =", i, ", yr =", yr, "\n")
    # yearly covariates
    covpred = brick(paste0("data/covars/covars_brick_", yr, "_20230316.grd"))
    covpred = covpred[[varnames]]
    pred = predict(covpred,
                   bestfits[[i]],
                   n.trees = bestfits[[i]]$gbm.call$best.trees,
                   type = "response")
    pred_year[[as.character(yr)]] = pred
  }
  saveRDS(pred_year, 
          paste0("outputs/pred_yearly_5var_occ_", ids[1], "_abs_", ids[2], "_" , tstamp(), ".rds"))
}
```

#### Calculation across samples (random occ and abs data sets)
quantile rasters are computed using raster::calc function
```{r}
fls = list.files("outputs/","^pred_yearly_5var_occ_([0-9]+)_abs_([0-9]+)_2023.*.rds$",full.names=TRUE)

nsamp = length(fls)
years = 2000:2020
probs = c(0.025,0.25,0.5,0.75,0.975)

library(raster)

predlist = vector("list", length=nsamp) # years 2000:2020
for (j in seq_len(nsamp)) {
  predlist[[j]] = readRDS(fls[j])
}

for (yr in years) {
  cat("yr =", yr, "\n")
  rstyrlist = vector("list", length=length(years))
  for (j in seq_len(nsamp)) {
    rstyrlist[[j]] = predlist[[j]][[as.character(yr)]]
  }
  
  rstcalc = calc(brick(rstyrlist), quantile,
                 probs=probs, na.rm=TRUE)
  saveRDS(rstcalc, 
          paste0("outputs/pred_5var_calc_", yr, "_" , tstamp(), ".rds"))
}

```


#### Probability of occurrence plots by year
Save as figures
```{r}
# plots
# pred_year = readRDS(paste0("outputs/pred_yearly_5var_20230320.rds"))
# fls = list.files("outputs/","^pred_5var_calc_([0-9]+)_20230321.rds",full.names=TRUE)
fls=list.files("outputs/","^pred_5var_calc_([0-9]+)_2023.*.rds$",full.names=TRUE)
years = 2000:2020

for (yr in years){
  cat("yr =", yr, "\n")
  id = grepl(paste0("calc_", yr), fls, fixed=TRUE)
  rst = readRDS(fls[id])
  p <- probocc_plot(raster=rst[[3]], theme=theme_map())
  rat <- 1
  ggsave(paste0("plots/probocc_", yr, "_",   tstamp(hour=T,minute=T), ".png"), p, width=7.4, height=7.4*rat, units="in")
}

## 2.5th, 25th, 50th, 75th, 97.5th percentile for 2020
yr = 2020
id = grepl(paste0("calc_", yr), fls, fixed=TRUE)
probs = c(0.025,0.25,0.5,0.75,0.975)
rst = readRDS(fls[id])

for (pr in seq_along(probs)) {
  plt <- probocc_plot(raster=rst[[pr]], theme=theme_map())
  rat <- 1
  ggsave(paste0("plots/probocc_", yr, "_", probs[pr],"_",   tstamp(hour=T,minute=T), ".png"), plt, width=7.4, height=7.4*rat, units="in")
}
```

#### Animation Probability of occurrence plots by year
Use the following command to create a gif file
Change the background, etc for a movie
It will be useful to create
1. change the background of the map to grey 
2. add year as the title

```{r}
# plots
# pred_year = readRDS(paste0("outputs/pred_yearly_5var_20230320.rds"))
# 
# years = 2000:2020
# for (yr in years){
#   cat("yr =", yr, "\n")
#   p <- probocc_plot(raster=pred_year[[as.character(yr)]], title=yr, theme=theme_ani())
#   rat <- 1
#   ggsave(paste0("plots/animation/probocc_", yr, "_",   tstamp(hour=T), ".png"), p, width=7.4, height=7.4*rat, units="in")
# }

fls = list.files("outputs/","^pred_5var_calc_([0-9]+)_2023.*.rds$",full.names=TRUE)
years = 2000:2020

for (yr in years){
  cat("yr =", yr, "\n")
  id = grepl(paste0("calc_", yr), fls, fixed=TRUE)
  rst = readRDS(fls[id])
  p <- probocc_plot(raster=rst[[3]], title=yr, theme=theme_ani())
  rat <- 1
  ggsave(paste0("plots/animation/probocc_", yr, "_",   tstamp(hour=T,minute=T), ".png"), p, width=7.4, height=7.4*rat, units="in")
}

```
convert -delay 100 -loop 0 -resize 50% plots/animation/*.png plots/animation/probocc_2000_2020.gif

convert -delay 100 -loop 0 plots/animation/*.png plots/animation/probocc_2000_2020.gif




#### Figure 4D: Probability of occurrence by region 
Identify regions with probability of occurrence with certain threshold
## Probability of occurrence by region
```{r}
# id <- find_best_fit(fits)
# rst <- brt_pred[[id]]

pred = readRDS("outputs/pred_5var_calc_2020_20230321.rds")
rst <- pred[[3]] # median value
rst_adm1 <- rst
rst_adm1[] <- NA
# af0 <- readRDS("data/africa_sub_Sahara_adm0_shp.rds")
af1 <- readRDS("data/africa_sub_Sahara_adm1_shp.rds")
af1_nona <- af1[!is.na(af1$NAME_1), ]
districts <- af1_nona$NAME_1

df <- data.frame(matrix(NA, nrow = length(districts), ncol = 2))
names(df) <- c("district", "prob")
df$district <- districts

# pop <- read_csv("outputs/")country_pop_20210524.csv
# ir <- read_csv("outputs/tf_inc_mean_20210524.csv")

# compute the mean values
for (i in 1:length(df$district)) {
  cat(i, ", ", df$district[i], "\n")
  poly = af1_nona[af1_nona$NAME_1 == df$district[i],]
  val = raster::extract(rst, poly, cellnumbers=T, df=T)
  meanval <-  mean(val[,3], na.rm=TRUE)
  df[i,2] <- meanval
  # af1_nona[af1_nona$NAME_1 == df$district[i], ] <- meanval
  rst_adm1[val$cell] <- meanval
}


p <- probocc_plot(raster=rst_adm1)
rat <- 1
ggsave(paste0("plots/probocc_adm1_", tstamp(), ".png"), p, width=7.4, height=7.4*rat, units="in")


```


### Aggregate at the regional and country level
```{r}
library(raster)
# load the prob of occ map 20 km
probnms <- c("q025","q250","q500","q750","q975")
# rstlist <- vector("list",length(nms))
list1 <- lapply(probnms, function(x) raster(paste0("outputs/",x, "_brt_pred_bkg_healthcare_full_covar20230224.tif")))

res1 <- aggregate_by_region(rst=list1[[1]], region="subregion", func="mean")

list2 <- lapply(list1, function(x) aggregate_by_region(rst=x, region="subregion", func="mean"))

# saveRDS(list2, "outputs/subregion_brt_pred_bkg_healthcare_full_covar20230224.rds")

dflist <- lapply(list2, function(x) x$data)

library(purrr)
library(dplyr)
df <- purrr::reduce(dflist, dplyr::left_join, by='area')
names(df) <- c("area", probnms)
write.csv(df, "outputs/subregion_brt_pred_bkg_healthcare_full_covar20230224.csv", row.names = F)


# aggregate at the country level
list3 <- lapply(list1, function(x) aggregate_by_region(rst=x, region="country", func="mean"))

saveRDS(list3, "outputs/country_brt_pred_bkg_healthcare_full_covar20230224.rds")

dflist <- lapply(list3, function(x) x$data)

library(purrr)
library(dplyr)
df <- purrr::reduce(dflist, dplyr::left_join, by='area')
names(df) <- c("area", probnms)
write.csv(df, "outputs/country_brt_pred_bkg_healthcare_full_covar20230224.csv", row.names = F)
```

### Aggregate at the regional and country level
Should I use the estimate, which is the median values out of $n$ simulations in which parameter values, background values, etc. vary
```{r}
library(raster)
# load the prob of occ map 20 km
probnms <- c("q500")
# rstlist <- vector("list",length(nms))
list1 <- lapply(probnms, function(x) raster(paste0("outputs/",x, "_brt_pred_bkg_healthcare_full_covar20230224.tif")))

res1 <- extract_grids_by_region(rst=list1[[1]], region="subregion")
# As an example grid cells for East Africa were extracted
# compute the mean
x = rst_poly$value[!is.na(rst_poly$value)]
set.seed(1)
meanx = mean(x)
## nonparameteric bootstraping CI
nrep = 100000
n = length(x);
meanx_re = replicate(nrep, mean(sample(x, n, rep=T)))
(quantile(meanx_re, c(0.025, 0.975)))
```

### Fig S8. ROC curve
```{r}
# plot(brt_fits[[1]])
# plot(brt_pred[[1]])
# fit_brt <- brt_fits[[1]]
occ_id=5
abs_id=5
yr=2000
predfls = list.files("outputs/","^pred_yearly_5var_occ_([0-9]+)_abs_([0-9]+)_2023.*.rds$",full.names=TRUE)
id = grepl(paste0("occ_", occ_id,"_abs_", abs_id), predfls, fixed=TRUE)
pred = readRDS(predfls[id])[[as.character(yr)]]

datafls = list.files("outputs/","^data_brt_occ_([0-9]+)_abs_([0-9]+)_2023.*.rds$",full.names=TRUE)
id = grepl(paste0("occ_", occ_id,"_abs_", abs_id), datafls, fixed=TRUE)
dat = readRDS(datafls[id])

nsamp = length(fls)
years = 2000:2020
probs = c(0.025,0.25,0.5,0.75,0.975)

fit_brt <- bestfit
# median prediction as the estimate 
calcpred = readRDS(paste0("outputs/calc_pred_half_5var_20230317T15.rds"))
pred_brt = calcpred$layer.3 # median raster
# fit_brt$cv.roc.matrix
# fit_brt$contributions
brtdat = readRDS(paste0("outputs/data_brt_20230316T184332.rds"))
occ_longlat = brtdat[bardat$yesno == 1, c("long","lat")] 
occ_longlat = brtdat[bardat$yesno == 1, c("long","lat")] 
p <- raster::extract(pred_brt, occ_longlat)
a <- raster::extract(pred_brt, abs_lonlat)
e <- evaluate(p=p, a=a)
# plot(e, 'ROC')
# e2 <- evaluate(p=occ_longlat, 
#                a=abs_lonlat, 
#                model=fit_brt,
#                x=covariates)
# plot(e2, "ROC")


tpr <- e@confusion[,"tp"]/(e@confusion[,"tp"] + e@confusion[,"fn"])
fpr <- e@confusion[,"fp"]/(e@confusion[,"fp"] + e@confusion[,"tn"])
prdf <- data.frame(tpr = tpr, fpr = fpr)

p <- ggplot(prdf, aes(fpr, tpr)) +
  geom_line() +
  theme_bw() +
  labs(x = "False positive rate", y = "True positive rate") 
p
ggsave(paste0("plots/roc_curve_",tstamp(), ".png"), p, width = 7.8*0.5, height = 8.2*0.5, units="in")
```

### Fig S7: 2.5th and 97.5th raster
```{r}
brtpred <- readRDS("outputs/brt_pred_bkg_healthcare_full_covar20230224.rds")
brt_pred <- brtpred[1:42]

# rst_calc <- raster::calc(raster::stack(brt_pred), function(x) {
#   quantile(x, probs = c(0.025, 0.25, 0.5, 0.75, 0.975), na.rm = TRUE)})
# # 
# probnms <- c("q025","q250","q500","q750","q975")
# for(i in 1:length(rst_calc)){
#   writeRaster(rst_calc[[i]], paste0("outputs/", probnms[[i]], "_brt_pred_bkg_healthcare_full_covar20230224.tif"), overwrite=T)
# } 

# saveRDS(rst_calc$layer.1, "outputs/q025_brt_pred_bkg_healthcare_full_covar20230224.rds")
# saveRDS(rst_calc$layer.2, "outputs/q250_brt_pred_bkg_healthcare_full_covar20230224.rds")
# saveRDS(rst_calc$layer.3, "outputs/q500_brt_pred_bkg_healthcare_full_covar20230224.rds")
# saveRDS(rst_calc$layer.4, "outputs/q750_brt_pred_bkg_healthcare_full_covar20230224.rds")
# saveRDS(rst_calc$layer.5, "outputs/q975_brt_pred_bkg_healthcare_full_covar20230224.rds")

rst_q2p5 <- probocc_plot(raster=rst_calc$layer.1)

ggsave(paste0("plots/prob_occ_q2p5_", tstamp(), ".png"), rst_q2p5, width=7.8*0.8, height=8.2*0.8, units="in")

rst_q50 <- probocc_plot(raster=rst_calc$layer.3)
ggsave(paste0("plots/prob_occ_q50_", tstamp(), ".png"), rst_q50, width=7.8*0.8, height=8.2*0.8, units="in")

rst_q97p5 <- probocc_plot(raster=rst_calc$layer.5)
ggsave(paste0("plots/prob_occ_q97p5_", tstamp(), ".png"), rst_q97p5, width=7.8*0.8, height=8.2*0.8, units="in")


# af@data$id <- rownames(af@data)
# af_points <- broom::tidy(af, region = "id")
# af_df <- dplyr::left_join(af_points, af@data, by = "id")

# # rst <- rst_calc$X2.5.
# rst <- rst_calc$layer.3
# names(rst) <- c("val")
# rst_pts <- raster::rasterToPoints(rst)
# rst_pts_df <- tbl_df(rst_pts)
# 
# p <- ggplot(rst_pts_df, aes(x = x, y = y, fill = val)) +
#   geom_raster() +
#   scale_fill_viridis("Probability \nof occurrence") + 
#   geom_polygon(data = af_df, aes(long, lat, group = group), 
#                fill = NA, inherit.aes = FALSE) +
#   geom_path(data = af_df, aes(long, lat, group = group), 
#             color = "black", inherit.aes = FALSE) +
#   coord_equal() + 
#   theme_map() 
# 
# ggsave(paste0("plots/prob_occ_q2p5_", tstamp(), ".png"), p, width=7.8*0.8, height=8.2*0.8, units="in")

```

### Fig 5. Plot relative influence
```{r}
library(dplyr)
library(ggplot2)
fits = readRDS("outputs/bestfits_20230321.rds")
d <- fits[[1]]$contributions
d$var
for (fit in fits) {
  d <- rbind(d, fit$contributions)
}
names(d) <- c("var", "rel_inf")
ds <- d %>% 
  group_by(var) %>% 
  summarize(avg=mean(rel_inf), sd=sd(rel_inf), 
            se=sd(rel_inf)/sqrt(n()), 
            q025=quantile(rel_inf, probs=0.025), 
            q975=quantile(rel_inf, probs=0.975),
            q500=quantile(rel_inf, probs=0.5),
            q250=quantile(rel_inf, probs=0.25), 
            q750=quantile(rel_inf, probs=0.75))
ds$var
ds$labels <- c("Improved water","Piped sanitation",
               "HIV","Underweight","Pf incidence rate")

plt <- ggplot(ds, aes(x=reorder(var,avg), y=q500))+
  geom_bar(stat="identity", fill="brown", alpha=0.6)+
  scale_x_discrete(labels=ds$labels[order(ds$avg)])+
  geom_errorbar(aes(ymin=q025, ymax=q975), width=0.3)+
  labs(y="Relative influence (%)", x=NULL) +
  coord_flip() + 
  theme_bw() + 
  theme(text=element_text(size=18),
        axis.text=element_text(size=18))
plt
ggsave(paste0("plots/prob_occ_rel_inf_", tstamp(), ".png"), plt, width = 7.7, height=7.3, units = "in")
```

### Fig 5. Variable response curve
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
# variable-response curve for each variable

# fits <- readRDS("outputs/brt_half_fit_20230317T045738.rds")
fits = readRDS("outputs/bestfits_20230321.rds")

n <- length(fits)
auc_cv <- data.frame(matrix(NA, nrow = n, ncol = 10))
var_res_df_list <- vector( "list", n)
resolution <- 100
nvar <- length(fits[[1]]$var.names)
for (i in 1:n) {
    auc_cv[i, ] <- fits[[i]]$cv.roc.matrix
    var_res_df <- data.frame(id = 1:resolution)
    for (j in 1:nvar) {
        d <- gbm::plot.gbm(fits[[i]], j, 
                           return.grid = TRUE, 
                           continuous.resolution = 100)
        var_res_df <- cbind(var_res_df, d)
    }
    var_res_df_list[[i]] <- var_res_df
}

var_res <- do.call('rbind', var_res_df_list)

# nm <- c("Water", "Sanitation", "Elevation", "HIV",
#         "Malaria", "Wasting")
# nm <- c("Pf incidence rate", "Pf parasite Rate", "Improved water", "Open defecation", "Piped sanitation", "Piped Water", "HIV", "Stunting", "Surface water", "Underweight", "Wasting")

nm <- c("Pf incidence rate", "Improved water", "Piped sanitation", "HIV",  "Underweight")

nm2 <- c("","_y")
nms <- unlist(lapply(nm, function(x) paste0(x,nm2)))

names(var_res) <- c("id", nms)

var <- var_res[,seq(2,(ncol(var_res)-1),2)]
res <- var_res[,seq(3,ncol(var_res),2)]

varlong <- pivot_longer(var, cols=1:ncol(var))
names(varlong) <- c("varname", "X") 
reslong <- pivot_longer(res, cols=1:ncol(res))

varreslong <- cbind(varlong, data.frame(Y=reslong$value))
varreslong$var <- factor(varreslong$var)
#log(p/(1-p)) = res
varreslong$Y <- exp(varreslong$Y)/(1 + exp(varreslong$Y))

plt <- ggplot(varreslong, aes(X, Y)) +
    geom_line(alpha = 0.2) +
    geom_smooth(fill = 'darkred', linewidth=0.3) +
    facet_wrap(~var, scales='free_x', strip.position='top') +
    labs(x="", y="Probability of occurrence" ) +
    theme(strip.background = element_blank(), strip.placement = "outside") +
    theme_bw() +
    scale_y_continuous(limits = c(0,1),
                       breaks = seq(0, 1, 0.25), 
                       labels = seq(0,1,0.25)) 
plt

ggsave(paste0("plots/var_res_curve_", tstamp(), ".png"),
       plt, width=3.4*2, height=2.7*2, units="in")
```


### Analyze brt fit results 2
```{r}
library(dplyr)
library(tidyr)
# variable-response curve for each variable
brt_fits = bestfits
n <- length(brt_fits)
auc_cv <- data.frame(matrix(NA, nrow = n, ncol = 10))
var_res_df_list <- vector( "list", n)
resolution <- 100
nvar <- length(brt_fits[[1]]$var.names)
for (i in 1:n) {
    auc_cv[i, ] <- brt_fits[[i]]$cv.roc.matrix
    var_res_df <- data.frame(id = 1:resolution)
    for (j in 1:nvar) {
        d <- gbm::plot.gbm(brt_fits[[i]], j, 
                           return.grid = TRUE, 
                           continuous.resolution = 100)
        var_res_df <- cbind(var_res_df, d)
    }
    var_res_df_list[[i]] <- var_res_df
}

var_res <- 
    data.frame(id = rep(1, length(var_res_df_list[[1]][, 2])), 
               Water = var_res_df_list[[1]][, 2],
               Water_y = var_res_df_list[[1]][, 3],
               Sanitation = var_res_df_list[[1]][, 4],
               Sanitation_y = var_res_df_list[[1]][, 5],
               Elevation = var_res_df_list[[1]][, 6],
               Elevation_y = var_res_df_list[[1]][, 7],
               HIV = var_res_df_list[[1]][, 8],
               HIV_y = var_res_df_list[[1]][, 9],
               Malaria = var_res_df_list[[1]][, 10],
               Malaria_y = var_res_df_list[[1]][, 11],
               Wasting = var_res_df_list[[1]][, 12],
               Wasting_y = var_res_df_list[[1]][, 13])


for (i in 2:n) {
  df <- data.frame(id = rep(1, length(var_res_df_list[[1]][, 2])),
                   Water = var_res_df_list[[i]][, 2],
               Water_y = var_res_df_list[[i]][, 3],
               Sanitation = var_res_df_list[[i]][, 4],
               Sanitation_y = var_res_df_list[[i]][, 5],
               Elevation = var_res_df_list[[i]][, 6],
               Elevation_y = var_res_df_list[[i]][, 7],
               HIV = var_res_df_list[[i]][, 8],
               HIV_y = var_res_df_list[[i]][, 9],
               Malaria = var_res_df_list[[i]][, 10],
               Malaria_y = var_res_df_list[[i]][, 11],
               Wasting = var_res_df_list[[i]][, 12],
               Wasting_y = var_res_df_list[[i]][, 13])
    
  var_res <- rbind(var_res, df)
}  

var <- var_res[, c(1, seq(2, 12, 2))] 
var_long <- gather(var, var, X, - id)

res <- var_res[, c(1, seq(3, 13, 2))] 
res_long <- gather(res, res, Y, -id)

var_res_long <- cbind(var_long, data.frame(Y = res_long$Y))
var_res_long$var <- factor(var_res_long$var)
#log( p/(1-p)) = res
var_res_long$Y <- exp(var_res_long$Y)/(1 + exp(var_res_long$Y))


p <- ggplot(var_res_long, aes(X, Y, group = id)) +
    geom_line(alpha = 0.2) +
    geom_smooth(fill = 'blue') +
    facet_wrap(~var, scales = 'free', strip.position = 'bottom') +
    labs(x=NULL, y = "Probability of occurrence" ) +
    theme(strip.background = element_blank(), strip.placement = "outside") +
    theme(axis.line = element_line() ) +
    scale_y_continuous(limits = c(0,1), breaks = seq(0, 1, 0.25), 
                       labels = seq(0,1,0.25)) 

    # annotate( "segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf ) +
    # annotate( "segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf )

# date <- format(Sys.Date(),"%Y%m%d")
# ggsave(paste0("plots/var_res_curve_", date, ".png"), p)
```




## Incidence rate by region
```{r}
rst <- readRDS("outputs/median_brt_pred_bkg_healthcare_20210616.rds")
af1 <- readRDS("data/africa_adm1_shp.rds")
af2 <- readRDS("data/africa_adm2_shp.rds")

af2_nona <- af2[!is.na(af2$NAME_2), ]
districts <- af2_nona$NAME_2

df <- data.frame(matrix(NA, nrow = length(districts), ncol = 2))
names(df) <- c("district", "prob")
df$district <- districts

# pop <- read_csv("outputs/")country_pop_20210524.csv
# ir <- read_csv("outputs/tf_inc_mean_20210524.csv")

for (i in 1:length(df$district)) {
  cat(i, ", ", df$district[i], "\n")
  poly <- af2_nona[af2_nona$NAME_2 == df$district[i], ]
  val <- unlist(raster::extract(rst, poly))
  df[i, 2] <- mean(val, na.rm = TRUE)
  
}

for (i in 1:length(df$district)) {
  cat(i, ", ", df$district[i], "\n")
  af2_nona[af2_nona$NAME_2 == df$district[i], ] <- df$prob[i]
}

df[is.nan(df[, 2]), 2] <- NA
af2$val <- NA   
af2[!is.na(af2$NAME_2), ]$val <- df$prob

polygon <- af2
if(!"id" %in% names(polygon)) {
  polygon$id <- 1:nrow(polygon)
}
# grid$fitted_pois <- mod3$summary.fitted.values$mean * grid$cellarea
poly_pts <- broom::tidy(polygon, region = "id")
poly_df <- plyr::join(poly_pts, polygon@data, by = "id")
library(ggplot2)
title = ""
xlab = ""
ylab = ""
title_size = 16
legend_title = ""
legend_position = c(0.2, 0.3)

ggplot(poly_df) +
  geom_polygon(aes(long, lat, group = group, fill = val), color = "black") +
  scale_color_discrete(legend_title) +
  labs(title = title, x = xlab, y = ylab) +
  theme(panel.background = element_blank(), # bg of the panel
        plot.background = element_blank(), # bg of the plot
        legend.background = element_blank(), # get rid of legend bg
        legend.box.background = element_blank(),
        panel.spacing = unit(c(0,0,0,0), "null"),
        plot.margin = unit(c(0,0,0,0), "null"),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = legend_position,
        plot.title = element_text(hjust = 0.5, size = title_size))
```


## Incidence rate by region
```{r}
rst <- readRDS("outputs/median_brt_pred_bkg_healthcare_20210616.rds")
# raster::calc function used to calculate median raster generates a temporary file
#  C:/Users/jonghoon.kim/AppData/Local/Temp/RtmpCa8V6J/raster/r_tmp_2021-12-13_133007_66132_27808.grd
# This seems to generate a problem, when the file was saved as rds and then 
# retrieved. 
# rst <- rst_calc$layer.3

af0 <- readRDS("outputs/africa_adm0_shp.rds")
if(!"id" %in% names(af0)) {
  af0$id <- 1:nrow(af0)
}
# grid$fitted_pois <- mod3$summary.fitted.values$mean * grid$cellarea
af0_pts <- broom::tidy(af0, region = "id")
af0_df <- plyr::join(af0_pts, af0@data, by = "id")
  
af1 <- readRDS("data/africa_adm1_shp.rds")
adm1 <- af1$NAME_1
af1$prob <- NA
df <- data.frame(matrix(NA, nrow = length(adm1), ncol = 2))
names(df) <- c("adm1", "prob")
df$adm1 <- adm1

# calculate the mean for the grid cells that fall on to a particular polygon
for (i in 1:length(adm1)) {
  cat(i, ", ", df$adm1[i], "\n")
  poly <- af1[af1$NAME_1 == df$adm1[i], ]
  val <- unlist(raster::extract(rst, poly))
  af1$prob[i] <- df[i, 2] <- mean(val, na.rm = TRUE)
}

af1_high <- af1
cutoff_prob <- 0.3
for (i in 1:length(af1_high$prob)){
  if(!is.na(af1_high$prob[i]) & af1_high$prob[i] < cutoff_prob){
    af1_high$prob[i] <- NA
  }
}
    
polygon <- af1_high

if(!"id" %in% names(polygon)) {
  polygon$id <- 1:nrow(polygon)
}
poly_pts <- broom::tidy(polygon, region = "id")
poly_df <- plyr::join(poly_pts, polygon@data, by = "id")

# to make country borders
adm0 <- unique(polygon$NAME_0)
polygon$id2 <- NA
for(i in seq_along(adm0)) {
  for(j in seq_along(polygon$NAME_0)) {
    if(polygon$NAME_0[j] == adm0[i]){
      polygon$id2[j] <- i
    }
  }
}
poly_pts_adm0 <- broom::tidy(polygon, region = "id2")
poly_df_adm0 <- plyr::join(poly_pts_adm0, polygon@data, by = "id")

# poly_pts_adm0_ <- broom::tidy(polygon, region = "id2")
# poly_pts_adm0_$id2 <- poly_pts_adm0_$id
# poly_pts_adm0_$id <- NULL
# poly_df_adm0_ <- plyr::join(poly_pts_adm0_, polygon@data, by = "id2")

# grid$fitted_pois <- mod3$summary.fitted.values$mean * grid$cellarea

library(ggplot2)
title = ""
xlab = ""
ylab = ""
title_size = 16
legend_title = "Probability\nof occurrence"
legend_position = c(0.2, 0.3)

library(dplyr)
library(viridis)
source("R/theme_map.R")
p <- 
  ggplot() +
  geom_polygon(data = poly_df, aes(long, lat, group = group, fill = prob), color = "black", size = 0.3) +
  geom_polygon(data = poly_df_adm0, aes(long, lat, group = group),
               fill = NA, color = "black", size = 0.8, inherit.aes = F)+
  # geom_polygon(data = af0_df, aes(long, lat, group = group),
  #              fill = NA, color = "black", size = 0.8, inherit.aes = F)+
  scale_fill_viridis(legend_title, na.value="white") + 
  labs(title = title, x = xlab, y = ylab) +
  theme_map(legend_position = legend_position) + 
  theme(plot.title = element_text(hjust = 0.5, size = title_size))

tstamp <- format(Sys.Date(), "%Y%m%d")
ggsave(paste("plots/iNTS_adm1_highprob2_", tstamp, ".png"), 
             p, width = 7.8*0.8, height = 8.2*0.8,
             units="in", dpi = 600)
```


## Probability of occurrence vs. TSAP incidence rate
```{r}
## median probability of occurrence


## Surveillance area
find_adm <- function(str, poly, adm = "all") {
  val <- NA
  if (adm == "all") {
    val <- grep(str, c(poly$NAME_1, poly$NAME_2, poly$NAME_3), value = TRUE)
  }
  else if (adm == 1) {
    val <- grep(str, c(poly$NAME_1), value = TRUE)
  } 
  else if (adm == 2){
    val <- grep(str, c(poly$NAME_2), value = TRUE)
  }
  return(val)
}

pol <- readRDS("data/gadm36_BFA_3_sp.rds")
find_adm("Nio", pol, adm = "all")



library(readxl)
dat <- read_xlsx("data/iNTS occurrence in sub-Saharan Africa since 2010.xlsx", sheet = "YES")
# extract TSAP
dat_ <- dat
dat <- dat[dat$STUDY_INFO == "Marks" & !is.na(dat$STUDY_INFO), c("PLACE_LOWEST_LEVEL", "COUNTRY",	"LOCATION",	"POLYGON",	"ADM", 	"LATITUDE",	"LONGITUDE")]

dat_inc <- read_xlsx("data/iNTS incidence systematic review final dataset.xlsx", sheet = "Final dataset")
dat_inc_ <- dat_inc
dat_inc <- dat_inc[dat_inc$`First author's last name` == "Marks" & dat_inc$`Publication year` == 2017 & !is.na(dat_inc$`First author's last name`), c("First author's last name","Study country", "Study locality (city, district, or province)", "Adjusted incidence")]

dat_inc$`Adjusted incidence`
dat_inc$all_ages <- c(237, 144, 37, rep(NA, 4), 7, 19, 32, 9, NA, 0)

dat_inc$PLACE <- dat_inc$`Study locality (city, district, or province)`
dat$PLACE <- dat$PLACE_LOWEST_LEVEL

dat_total <- left_join(dat, dat_inc, by = "PLACE")
dat_total_ <- dat_total
dat_total <- dat_total[, c("COUNTRY", "POLYGON", "ADM", "LATITUDE", "LONGITUDE", "PLACE",  "Adjusted incidence", "all_ages")]

df <- dat_total
df_temp <- df
for (i in 1:nrow(df)) {
  if(df_temp$POLYGON[i] == "NA"){
    df$POLYGON[i] <- NA
  }
}

# prob_rst <- readRDS("outputs/median_brt_pred_bkg_healthcare_full_covar20210627.rds")
prob_rst <- readRDS("outputs/median_brt_pred_bkg_healthcare_20210616.rds")

library(raster)
df$prob <- NA
for (i in 1:nrow(df)) {
  # cat(i, ", ", "\n")
  if(is.na(df$POLYGON[i])) {
    df$prob[i] <- 
      raster::extract(prob_rst, matrix(c(df$LONGITUDE[i], df$LATITUDE[i]), nrow = 1, ncol = 2))
  } else {
    cntry <- af2[af2$NAME_0 == df$COUNTRY[i],]
    # pol <- NULL
    if (df$ADM[i] == 1) {
      pol <- cntry[cntry$NAME_1 == df$POLYGON[i],]
    } else if(df$ADM[i] == 2) {
      pol <- cntry[cntry$NAME_2 == df$POLYGON[i],]
    }
    vals <- raster::extract(prob_rst, pol)
    df$prob[i] <- mean(unlist(vals))
  }
}

# saveRDS(dat_total, "outputs/dat_total.RDS")
library("Hmisc")
df2 <- df[!is.na(df$all_ages) & !is.na(df$prob), ]
rcorr(as.matrix(df2[, c("prob", "all_ages")]))
rcorr(as.matrix(df2[3:8, c("prob", "all_ages")]))
```


## Sensitivity analysis 
### Different k for the exponential distribution
We assumed that the 1/20 and 1/100 as the baseline scenario and we varies k for the 0.005 allowing patients from longer distance
```{r}
# yearly covariate sets
# 20km, admin level 2, 1, 0, and African subregion, and overall
```

Two rasters (XX and ir_*) have the same resolutions and were not aligned perfectly and therefore, resample is done as ppp for 0-1 yo as the reference


### Check polygons
```{r}
shplist <- readRDS("outputs/occ_shapelist_20230317.rds")
for(i in 1:length(shplist)){
  # print(placelist[[i]])
  cat("i =", i, ", name =", shplist[[i]]$name, "\n")
  for(n in 1:4) {
    print(eval(parse(text=paste0("shplist[[i]]$shape$NAME_", n))))
  }
  readline(prompt="\n-----------------------------")
}
```
Identify simulation ids that were already conducted
Combine the data from separate simulations
```{r}
fls = list.files("outputs/", pattern="brt_fit_20230316.*", full.names=T)
ids = c()
fits = vector('list', 20)
for (fl in fls){
  lst = readRDS(fl)
  for (i in 1:length(lst)) {
    if (!is.null(lst[[i]])) {
      ids = c(ids, i)
      fits[[i]] = lst[[i]]
    }
  }
}
saveRDS(fits, paste0("outputs/brt_fits_", tstamp(), ".rds"))
```

### Raster calculation (median raster)
```{r}
brt_pred <- readRDS("outputs/bpred_half_seed_5var_20230317T152221.rds")
rst_calc <- raster::calc(stack(brt_pred), function(x) {
  quantile(x, probs = c(0.025, 0.25, 0.5, 0.75, 0.975), na.rm = TRUE)})

saveRDS(rst_calc, paste0("outputs/calc_pred_half_5var_", tstamp(hour=T), ".rds"))
```
### Random variation

for a given set of hyper-parameters that produced the best fit, we conducte a number of simulations to examine variations
```{r}
# identify the best fit

# to extract covariates used in the modeling
varnames = readRDS("outputs/name_cov_baseline.rds")
varnames = c("X202206_Global_Pf_Incidence_Rate",
             "improved_water_20km",
             "piped_sanitation_20km",
             "prev_HIV_adults_20km","underweight_20km")   
# identify the best fit
fits = readRDS("outputs/brt_half_fit_5_var_20230317T093744.rds")
id <- find_best_fit(fits)
bestfit = fits[[id]]

niter = 200
yr = 2020
pred_2020_seed = vector("list", niter)
names(pred_2020_seed) = paste0("seed", 1:niter)
# predict for 

for (i in seq_len(niter)) {
  set.seed(i)
  cat("i =", i, "\n")
  covpred = brick(paste0("data/covars/covars_brick_", yr, "_20230316.grd"))
  covpred = covpred[[varnames]]
  pred_2020_seed[[i]]=predict(covpred,
                 bestfit,
                 n.trees=bestfit$gbm.call$best.trees,
                 type="response")
}

saveRDS(pred_2020_seed, paste0("outputs/pred_half_seed_5var_", tstamp(hour=T,minute=T,second=T), ".rds"))

```

Identify the regions 
```{r}
id <- find_best_fit(fits)
rst <- brt_pred[[id]]
rst_02 <- rst
rst_02[rst_02 < 0.2] <- NA

p <-
rat <- 1
ggsave(paste0("plots/probocc_", tstamp(), ".png"), p, width=7.4, height=7.4*rat, units="in")
```

