---
title: "Geospatial distribution of invasive non-typhoidal Salmonella (iNTS) disease occurrence in sub-Saharan Africa"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overall structure

This document is about the paper titled **Geospatial distribution of invasive non-typhoidal Salmonella (iNTS) disease occurrence in sub-Saharan Africa**. It describes the procedure of creating a random data set based on a systematic literature review. 

 
### 1. Creating random occurrence and background data

You can include R code in the document as follows:

```{r cars}
library(raster)
library(dplyr)
# occ = readRDS("outputs/occ_dataset_20230314T1231.rds")
# occfull = readRDS("outputs/occ_dataset_20230316T173539.rds")
# occfull = readRDS("outputs/occ_dataset_20230317T111754.rds")
occfull = readRDS("outputs/occ_dataset_dist3_200_20230318.rds")
refrst = raster("data/covars/motorized_travel_time_to_healthcare_20km_af_20230314.tif")
# go through data year by year such that we match the covariates by year and location
for (i in 15) {
  for (j in 16:20) {
    cat("i =", i, ", j =",  j, "\n")
    set.seed(j)
    data_brt = create_occ_abs_data(occdata=occfull[[i]], refraster=refrst)
    saveRDS(data_brt, paste0("outputs/data_dist3_200_brt_occ_", i, "_abs_", j, "_", tstamp(),".rds"))
  }
}
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
